Memes: a motif analysis environment in R using tools from the MEME Suite

### 2021-07-13 manuscript.Rmd
fix title & table callouts

---
title: 'Memes: an R interface to the MEME Suite'
author:
- ORCID: 0000-0003-1000-1579
  name: Spencer L. Nystrom
- ORCID: 0000-0001-8226-0604
  name: Daniel J. McKay
output: 
  bookdown::word_document2:
    fig_caption: yes
    reference_docx: template.docx
    keep_tex: true

@@

Here we present `memes`, a motif analysis package that provides a seamless
interface to a selection of popular MEME Suite tools within R. Six MEME Suite tools are contained within the `memes` wrapper (Table 1). `memes` uses base R and Bioconductor
data types for data input and output, facilitating better integration with
common analysis tools like the `tidyverse` and other Bioconductor packages [@rcore; @bioconductor; @tidyverse].
Unlike the commandline implementation, `memes` outputs also function as inputs
to other MEME Suite tools, allowing simple construction of motif analysis
pipelines without additional data processing steps. Additionally, R/Bioconductor
data structures provide a full-featured representation of MEME Suite output
data, providing users quick access to all relevant data structures with simple
syntax.

@@

Using this approach, the results can be visualized using  the
`view_tomtom_hits()` function to visually inspect the matches assigned by
TomTom, providing a simple way to assess the quality of the match.  A
representative plot of these data is shown in **Figure \@ref(fig:tomtom-example)**, while
the full results can be viewed in **Table 1**. This analysis revealed
that the motifs associated with decreasing sites are highly similar to the E93
motif. Conversely, motifs found in increasing sites match the transcription
factors Br, L(3)neo38, Gl, and Lola (**Table 1**). These data support
the hypothesis that E93 binding to its motif may play a larger role in chromatin
closing activity, while binding of the transcription factor br is associated
with chromatin opening.

>>>>

---
title: 'Memes: a motif analysis environment in R using tools from the MEME Suite'
author:
- ORCID: 0000-0003-1000-1579
  name: Spencer L. Nystrom
- ORCID: 0000-0001-8226-0604
  name: Daniel J. McKay
output: 
  bookdown::word_document2:
    fig_caption: yes
    reference_docx: template.docx
    keep_tex: true

@@

Here we present `memes`, a motif analysis package that provides a seamless
interface to a selection of popular MEME Suite tools within R. Six MEME Suite tools are contained within the `memes` wrapper (**Table 1**). `memes` uses base R and Bioconductor
data types for data input and output, facilitating better integration with
common analysis tools like the `tidyverse` and other Bioconductor packages [@rcore; @bioconductor; @tidyverse].
Unlike the commandline implementation, `memes` outputs also function as inputs
to other MEME Suite tools, allowing simple construction of motif analysis
pipelines without additional data processing steps. Additionally, R/Bioconductor
data structures provide a full-featured representation of MEME Suite output
data, providing users quick access to all relevant data structures with simple
syntax.

@@

Using this approach, the results can be visualized using  the
`view_tomtom_hits()` function to visually inspect the matches assigned by
TomTom, providing a simple way to assess the quality of the match.  A
representative plot of these data is shown in **Figure \@ref(fig:tomtom-example)**, while
the full results can be viewed in **Table 2**. This analysis revealed
that the motifs associated with decreasing sites are highly similar to the E93
motif. Conversely, motifs found in increasing sites match the transcription
factors Br, L(3)neo38, Gl, and Lola (**Table 2**). These data support
the hypothesis that E93 binding to its motif may play a larger role in chromatin
closing activity, while binding of the transcription factor br is associated
with chromatin opening.
